---
output:
  pdf_document: default
  html_document: default
---

# Students' Development of Statistical Narratives in Probability Modeling SEM

### Issues So Far

-   Total N = **84**
-   Complete cases = **40**
-   A lot of missing observations in NE4 (30 missing) and NE5 (27 missing), casuing the full SEM to become unstable due to small sample size

```{r}
library(lavaan) # for SEM
library(tidyverse, quietly = TRUE) # for data set join, file link read
```

```{r}
# get data from all three activities
A3 = read_csv("https://raw.githubusercontent.com/jfleege/narrative-sem/main/datasets/CombinedActivity3.csv", show_col_types = FALSE)%>%
  rename(Anonymous.ID = `Anonymous ID`)

A3$Survey_NE = as.numeric(A3$Survey_NE)

A4 = read_csv("https://raw.githubusercontent.com/jfleege/narrative-sem/main/datasets/CombinedActivity4.csv", show_col_types = FALSE)%>%
  rename(Anonymous.ID = `Anonymous ID`)

A4$Survey_NE = as.numeric(A4$Survey_NE)

A5 = read_csv("https://raw.githubusercontent.com/jfleege/narrative-sem/main/datasets/CombinedActivity5.csv", show_col_types = FALSE) %>%
  rename(Anonymous.ID = `Anonymous ID`)

A5$Survey_NE = as.numeric(A5$Survey_NE)

# establish ANIQ columns
item_cols <- c("AW1","AW2","AW3","AW4","AW5",
               "TM1","TM2","TM3","TM4","TM5",
               "CA1","CA2","CA3","CA4","CA5",
               "TH1","TH2","TH3","TH4","TH5")

# combine data across all three activities
SEMcombined <- A3 %>%
  select(Anonymous.ID, all_of(item_cols), NE3 = Survey_NE) %>%
  left_join(
    A4 %>% select(Anonymous.ID, NE4 = Survey_NE),
    by = "Anonymous.ID"
  ) %>%
  left_join(
    A5 %>% select(Anonymous.ID, NE5 = Survey_NE),
    by = "Anonymous.ID"
  )

# convert ANIQ item columns to numeric
SEMcombined[item_cols] <- lapply(
  SEMcombined[item_cols],
  function(x) as.numeric(as.character(x))
)

# order data
SEMcombined$NE3 <- as.ordered(SEMcombined$NE3)
SEMcombined$NE4 <- as.ordered(SEMcombined$NE4)
SEMcombined$NE5 <- as.ordered(SEMcombined$NE5)
```

### Data Set Information

```{r}
sum(complete.cases(SEMcombined)) # complete cases
nrow(SEMcombined) # total number of observations (students)
colSums(is.na(SEMcombined)) # number of missing values per column
```

I thought it might be helpful to include this brief information about the data set before diving into the models to get a better understanding of why we might be encountering certain problems. After running these commands, I found that there are 40 complete cases out of 84 total observations (probably not good), with some missing values in the ANIQ item columns.

### Model 1: Baseline

This was the first structure made and tested, with no changes to any of the ANIQ columns or subcategories. Not favored due to the bad scaled CFI, RMSEA, and SRMR values of 0.622, 0.549, and 0.123, respectively. Initial conclusion for the bad fit was due to the lack of overall used observations (n = 40) given the amount of ANIQ parameters.

```{r}
model1 <- "
  aw =~ AW1 + AW2 + AW3 + AW4 + AW5
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA1 + CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH2 + TH3 + TH4 + TH5

  NE3 ~ aw + tm + ca + th
  NE4 ~ aw + tm + ca + th
  NE5 ~ aw + tm + ca + th
"

fit1 <- sem(model1, data = SEMcombined, ordered = c("NE3", "NE4", "NE5"))

summary(fit1, fit.measures = TRUE, standardized = TRUE)
```

### Model 2: EFA Adjustments

This model was similar to the first; however, removed the ANIQ items that either had low factor loadings or cross-loadings based on Jon's findings. Based on these adjustments, noticed that CFI improved slightly to 0.662 and TLI improved to 0.577, but SRMR fell to 0.110 on n = 40 observations.

```{r}
model2 <- "
  aw =~ AW1 + AW2 + AW3 + AW4
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH3 + TH4 + TH5

  NE3 ~ aw + tm + ca + th
  NE4 ~ aw + tm + ca + th
  NE5 ~ aw + tm + ca + th
"

fit2 <- sem(model2, data = SEMcombined, ordered = c("NE3", "NE4", "NE5"))

summary(fit2, fit.measures = TRUE, standardized = TRUE)
```

### Model 3: Activity Dependency

This model was made considering that the latter activities (4 and 5) were dependent on the previous as the students progressed through the activities. Additionally, this model uses the same subcategories as Model 2. This model had the best fit of the three with a CFI of 0.967, TLI of 0.959, and SRMR of 0.082. Unsure if this is due to over fitting.

```{r}
model3 <- "
  aw =~ AW1 + AW2 + AW3 + AW4
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH3 + TH4 + TH5

  NE3 ~ aw + tm + ca + th
  NE4 ~ NE3 + aw + tm + ca + th
  NE5 ~ NE4 + aw + tm + ca + th
"

fit3 <- sem(
  model3,
  data = SEMcombined,
  ordered = c(item_cols, "NE3","NE4","NE5"),
  estimator = "WLSMV",
  parameterization = "theta",
  missing = "pairwise"
)

summary(fit3, fit.measures = TRUE, standardized = TRUE)
```

### Conclusion

Looking for guidance on how to appropriately model this data set given the number of missing observations and whether a reduced SEM, CFA, or maybe an alternative approach is most appropriate for our situation.

```{r}
model4 <- "
  aw =~ AW1 + AW2 + AW3 + AW4
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH3 + TH4 + TH5

  NE5 ~ NE3 + NE4 + aw + tm + ca + th
"

fit4 <- sem(
  model4,
  data = SEMcombined,
  ordered = c("NE3","NE4","NE5"),
  estimator = "WLSMV",
  parameterization = "theta",
  missing = "pairwise"
)

summary(fit4, fit.measures = TRUE, standardized = TRUE)
```

```{r}
model5 <- "
  aw =~ AW1 + AW2 + AW3 + AW4
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH3 + TH4 + TH5
  
  NI =~ aw + tm + ca + th
  NE3 ~~ NE4
  NE3 ~~ NI
  NE4 ~~ NI
  NE5 ~ NE3 + NE4 + NI
"

fit5 <- sem(
  model5,
  data = SEMcombined,
  ordered = c("NE3","NE4","NE5"),
  estimator = "WLSMV",
  parameterization = "theta",
  missing = "pairwise"
)

summary(fit5, fit.measures = TRUE, standardized = TRUE)
```

```{r}
model6 <- "
  aw =~ AW1 + AW2 + AW3 + AW4
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH3 + TH4 + TH5
  
  NI =~ aw + tm + ca + th
  NE4 ~~ NI
  NE5 ~ NE4 + NI
"

fit6 <- sem(
  model6,
  data = SEMcombined,
  ordered = c("NE4","NE5"),
  estimator = "WLSMV",
  parameterization = "theta",
  missing = "pairwise"
)

summary(fit6, fit.measures = TRUE, standardized = TRUE)
```

```{r}
model7 <- "
  aw =~ AW1 + AW2 + AW3 + AW4
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH3 + TH4 + TH5
  
  NE4 ~~ aw + tm + ca + th
  NE5 ~ NE4 + aw + tm + ca + th
"

fit7 <- sem(
  model7,
  data = SEMcombined,
  ordered = c("NE4","NE5"),
  estimator = "WLSMV",
  parameterization = "theta",
  missing = "pairwise"
)

summary(fit7, fit.measures = TRUE, standardized = TRUE)
```

```{r}
library(car)

SEMcombined$aw <- rowMeans(SEMcombined[, c("AW1","AW2","AW3","AW4")], na.rm = TRUE)
SEMcombined$tm <- rowMeans(SEMcombined[, c("TM1","TM2","TM3","TM4","TM5")], na.rm = TRUE)
SEMcombined$ca <- rowMeans(SEMcombined[, c("CA2","CA3","CA4","CA5")], na.rm = TRUE)
SEMcombined$th <- rowMeans(SEMcombined[, c("TH1","TH3","TH4","TH5")], na.rm = TRUE)

A3_log <- glm(NE3 ~ aw + tm + ca + th, data = SEMcombined, family = binomial)
A4_log <- glm(NE4 ~ aw + tm + ca + th, data = SEMcombined, family = binomial)

model8 <- "
  aw =~ AW1 + AW2 + AW3 + AW4
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH3 + TH4 + TH5

  NE3 ~ aw + tm + ca + th
  NE4 ~ NE3 + aw + tm + ca + th
  NE5 ~ NE4 + aw + tm + ca + th
"

fit8 <- sem(
  model8,
  data = SEMcombined,
  ordered = c("NE3", "NE4", "NE5"),
  estimator = "WLSMV",
  parameterization = "theta",
  missing = "pairwise"
)

summary(fit8, fit.measures = TRUE, standardized = TRUE)
```

```{r}
model9 <- "
  aw =~ AW1 + AW2 + AW3 + AW4
  tm =~ TM1 + TM2 + TM3 + TM4 + TM5
  ca =~ CA2 + CA3 + CA4 + CA5
  th =~ TH1 + TH3 + TH4 + TH5

  NE3 ~ aw + tm + ca + th
  NE4 ~ aw + tm + ca + th
  NE5 ~ NE3 + NE4 + aw + tm + ca + th
"

fit9 <- sem(
  model9,
  data = SEMcombined,
  ordered = c("NE3", "NE4", "NE5"),
  estimator = "WLSMV",
  parameterization = "theta",
  missing = "pairwise"
)

summary(fit9, fit.measures = TRUE, standardized = TRUE)
```

